### 2012-02-10 GUIErrorDetector/paper/evaluation.tex
more small edits

For all \subnum programs, we manually determined validity of each output warning
by either searching the
bug repository to check whether the same error has been reported before,
or submitting a new bug report, or writing a test driver to reproduce
the error.

\begin{table*}[ht]
\begin{center}
 \fontsize{9pt}{\baselineskip}\selectfont
\hspace*{-0.2cm}
\setlength{\tabcolsep}{.15\tabcolsep}
\begin{tabular}{|l||c|c|c||c|c|c||c|c|c||c|}
\hline
 Subject&  \multicolumn{9}{|c||}{Our Technique} & Requiring Wrappers  \\
\cline{2-10}
 Program  &  \multicolumn{3}{|c||}{RTA }& \multicolumn{3}{|c||}{0-CFA } & \multicolumn{3}{|c||}{1-CFA } & (Section~\ref{sec:straightforward})  \\
\cline{2-11}
 & CG Size & \#Warning & \#Bug & CG Size & \#Warning & \#Bug & CG Size & \#Warning & \#Bug & \#Warning \\
\hline \hline
\multicolumn{11}{|l|}{SWT desktop applications}   \\
 \hline
 VirgoFTP&  12401 &  1 &  1 & 10858 & 1 & 1 & 43598 & 2 & 2& 149  \\
 \hline
 FileBunker &  18951 &  1 &  0 & 15743 & 0 & 0 & 76088 & 2 & 1& 693  \\
 \hline
 ArecaBackup&  20882 &  1  &  0 & 19697 & 1 & 1  & 116398 & 1 & 1 & 3021\\
 \hline
 \hline
\multicolumn{11}{|l|}{Eclipse plugins}   \\
 \hline
 EclipseRunner&  11248&  6 &  1 & 7201 & 6 & 1 & 26911 & 6 & 1& 202  \\
 %\hline
 %FileSync&  12132 &  xx &  xxx &8235 & xx & xx& 32565 & 1 & 1 & 331 & 1 \\
 \hline
 HudsonEclipse& 18473 &  2 &  1 & 15814 & 2 & 1& 56645 & 3 & 1 & 182 \\
 \hline
 \hline
\multicolumn{11}{|l|}{Swing desktop applications}   \\
 \hline
 S3dropbox & 37751 &  0 &  0 & 30609 & 0 & 0 & 115324 & 1 & 1 & 210  \\
 \hline
  SudokuSolver&  27730&  3 &  2 & 20907 & 3 & 2 & 39299 & 2 & 2 & 356  \\
 \hline
 \hline
\multicolumn{11}{|l|}{Android mobile applications}   \\
 \hline
 SGTPuzzler & 13631 / 13865&  1 / 16 &  0 / 0 & 9546 / 9682& 0 / 4& 0 / 1 & 35198 / 35756 & 0 / 1  & 0 / 1& 104 \\
 \hline
 Fennec & 14058 / 14387 &  1 / 1 &  0 / 0 & 8263 / 8898 & 1 / 1 & 0 / 0& 29125/ 31759 & 3 / 3 & 1 / 1& 433 \\
 \hline
 MyTracks & 24036 / 24036 &  161 / 220 & 0 / 0 & 10803 / 13645 & 119 / 119 & 0 / 0 & 39235 / 110977 & 1 / 1 & 0 / 1 & 1192 \\
\hline
\hline
 Total & 199161 / 199724 &  177 / 251 & 5 / 5 & 149441 / 153054& 133 / 137& 6 / 7 & 577821 / 652756& 21 / 22 & 10 / 12 & 6542\\
\hline
\end{tabular}
\end{center}
\vspace{-15pt}
\Caption{{\label{table:results}Experimental results in finding invalid thread access errors
in multithreaded GUI programs. Column ``CG Size'' shows the
call graph size in terms of node number. Column ``\#Warning''
shows the number of warnings issued by our tool. Column ``\#Bug'' shows
the actual bugs found. Columns ``RTA '', ``0-CFA'',
and ``1-CFA'' show the results of using different
call graph construction algorithms. In each cell of the table
for Android applications, a slash ``/'' separates the result of
using standard call graph construction algorithm and the reflection-aware
call graph construction algorithm (with \annotationnum native method annotations
for SGTPuzzler). The \newbugs errors found in VirgoFTP, FileBunker, ArecaBackup
S3dropbox, and SudokuSolver are previously  unknown. As a comparison, the
results of the ``Requiring Wrappers'' approach (Section~\ref{sec:straightforward}),
are shown at the far right.} }
\end{table*}

@@

We submitted all \newbugs new errors
to the subject developers. As of February 2012, 1 error in S3dropbox has been confirmed,
4 errors have been reproduced and validated in our own experimental environment, and
2 errors in FileBunker and ArecaBackup had not yet been verified by developers.

@@

%Figures~\ref{fig:swingerror} amd~\ref{fig:pluginerror} show
%two real errors our tool has found. 
Figure~\ref{fig:swingerror}
shows an invalid thread access error our tool found in S3dropbox.
This error happens when the \CodeIn{deleteCurrentBucket}
call on line 63 throws an exception, making it hard to
detect by testing. We reported this error to the S3dropbox developers. Tom Czarniecki,
a key developer of S3dropbox, confirmed this single-GUI-thread
violation. He mentioned that the S3dropbox project uses
certain design patterns to avoid such violations (e.g.,
actions for UI interaction are encapsulated into a \CodeIn{Worker} interface),
but the developers still overlooked  the error our tool found.
Another reason they overlooked this violation is because some GUI frameworks like Swing
do not provide any run-time checks for invalid thread accesses. The Swing
GUI does not exhibit user-visible faults on some erreoneous executions.
However, as clearly stated in the official documentation~\cite{swing}, %Swing is a single-GUI-threaded GUI toolkit and
accessing Swing GUI objects from non-UI threads risks thread interference
or memory-consistency errors.

@@

Figure~\ref{fig:pluginerror} shows an error found in the EclipseRunner
plugin. This error is event-related. It happens when a 
\CodeIn{ResourceChangeEvent} happen, which then invokes the \CodeIn{refresh}
method on line 415 to update the user interface. In EclipseRunner,
the \CodeIn{refresh} method is
called by 6 different methods from the same non-UI thread. Thus, our
tool issues 6 separate warnings to indicate
6 different ways to trigger this error. 5 of the warnings are redundant.

Besides the above two examples, other errors our tool reported are also
subtle to find. For example, our tool found two new errors in SudokuSolver.
One error only happens when
the given Sudoku is unsolvable; and the other one happens
when  JVM fails to launch the %mail composing window of the
user default mail client.% (after a user clicks the "eMail Me" button).

As reflected in our experiments, we found GUI developers have already
used design patterns, runtime checks, and testing to avoid violating
the single-GUI-thread rule. However, due to the huge space of
possible UI interactions, potential
invalid thread access errors still exist in many corner cases and
can be hard to find. 
%Thus, a static analysis as presented in this paper
%would be useful.

@@

We next compare the three call graph construction algorithms (RTA, 0-CFA, and 1-CFA)
used in our experiments. As shown in Table~\ref{table:results},  1-CFA found
more errors than the other two algorithms. This
is because RTA and 0-CFA do not consider the calling context when
constructing a call graph, and may
mix the same method calls by different callers into a single node, thus
introducing imprecision. Although the error-revealing paths exist in
less precise graphs, the impreicison prevents our algorithm from finding them.
Figure~\ref{fig:ex} illustrates this point.

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.45]{cgexample}
  \vspace*{-5.0ex}\caption {{\label{fig:ex} (a) shows an example code, in which
\CodeIn{safe()} is a Safe UI method, and \CodeIn{uiAcc()} is a UI-Accessing method.
(b) shows a less precise call graph built by RTA or 0-CFA, and (c) shows a more
precise call graph built by 1-CFA. Nodes for constructors are omitted for brevity.
Using the less precise call graph, the error
detection algorithm in Figure~\ref{fig:detectalgorithm} will report an invalid method call chain: \CodeIn{entry()}
$\rightarrow$ \CodeIn{f()} $\rightarrow$ \CodeIn{start()} $\rightarrow$ \CodeIn{run()}
$\rightarrow$ \CodeIn{uiAcc()},
but miss the actual error path: \CodeIn{entry()} $\rightarrow$ \CodeIn{g1()} $\rightarrow$
\CodeIn{g2()} $\rightarrow$ \CodeIn{start()} $\rightarrow$ \CodeIn{run()} $\rightarrow$ \CodeIn{uiAcc()}. The
actual error path can only be found when using the more precise call graph.
}}
\end{figure}

@@

The first variant uses DFS for graph traversal by changing
the data structure of $\mathit{worklist}$ on line 7 in Figure~\ref{fig:detectalgorithm} to
a stack, and changing queue operations (\CodeIn{dequeue}, and \CodeIn{enqueueAll}
on lines 10, and 21 in Figure~\ref{fig:detectalgorithm}, respectively)
to the corresponding statck operations (\CodeIn{pop}, and \CodeIn{pushAll}).
This variant found 11 errors in total and missed 1 error in the FileBunker subject.
The reason is that DFS tends to search deeper into the graph, and the path it returns
may not be the shortest one and is more likely to be an infeasible path in practice.
%For the FileBunker subject, paths
%returned by DFS are much longer than BFS, and do not actually exist.

@@

The second variant uses exhaustive search to find potential errors. This variant
enumerates all non-cyclic paths from all reachable \CodeIn{Thread.start()} nodes
to each UI-accessing node, and then checks whether each
path spawns a new thread and accesses GUI objects without using safe UI methods.
This variant ran out of memory on all subjects. We checked
the explored paths before the algorithm terminated, and found none of them
revealed a real error. This is because the number of
non-cyclic paths in a graph is exponential in the graph size.%node number.
For a realistic call graph which often contains thousands of nodes,
it is infeasible to enumerate all paths.
%only a very few amount of paths can be enumerated within a practical amount of time.

@@

\noindent \textbf{\textit{Limitations.}}
Our technique is limited in three aspects. First, it only considers
non-UI threads that are spawned by the UI-thread after the GUI
is initialized, and ignores other possible non-UI threads (quite unusual)
that are created during the pre-initialization GUI work. One way
to remedy this limitation is to design an analysis to identify
those non-UI threads created before a GUI is launched.
Second, like many bug-finding techniques, our technique
is neither sound nor complete. It may miss true positives due
to the used graph search strategy, or
issue false positives due to the conservative nature of
a static analysis. Designing better call graph construction algorithms
and graph search strategies may alleviate this limitation.
Third, our tool cannot compute call relationships
for inter-process communication between components. It also
requires users to manually add annotations to characterize
call relationships that involve native methods. This limitation
may lead to false negatives. Investigating the false negative
rate is ongoing work.

>>>>

For all \subnum programs, we manually determined the validity of each output warning
by either searching the
bug repository to check whether the same error has been reported before,
or submitting a new bug report, or writing a test driver to reproduce
the error.

\begin{table*}[ht]
\begin{center}
 \fontsize{9pt}{\baselineskip}\selectfont
\hspace*{-0.2cm}
\setlength{\tabcolsep}{.15\tabcolsep}
\begin{tabular}{|l||c|c|c||c|c|c||c|c|c||c|}
\hline
 Subject&  \multicolumn{9}{|c||}{Our Technique} & Requiring Wrappers  \\
\cline{2-10}
 Program  &  \multicolumn{3}{|c||}{RTA }& \multicolumn{3}{|c||}{0-CFA } & \multicolumn{3}{|c||}{1-CFA } & (Section~\ref{sec:straightforward})  \\
\cline{2-11}
 & CG Size & \#Warning & \#Bug & CG Size & \#Warning & \#Bug & CG Size & \#Warning & \#Bug & \#Warning \\
\hline \hline
\multicolumn{11}{|l|}{SWT desktop applications}   \\
 \hline
 VirgoFTP&  12401 &  1 &  1 & 10858 & 1 & 1 & 43598 & 2 & 2& 149  \\
 \hline
 FileBunker &  18951 &  1 &  0 & 15743 & 0 & 0 & 76088 & 2 & 1& 693  \\
 \hline
 ArecaBackup&  20882 &  1  &  0 & 19697 & 1 & 1  & 116398 & 1 & 1 & 3021\\
 \hline
 \hline
\multicolumn{11}{|l|}{Eclipse plugins}   \\
 \hline
 EclipseRunner&  11248&  6 &  1 & 7201 & 6 & 1 & 26911 & 6 & 1& 202  \\
 %\hline
 %FileSync&  12132 &  xx &  xxx &8235 & xx & xx& 32565 & 1 & 1 & 331 & 1 \\
 \hline
 HudsonEclipse& 18473 &  2 &  1 & 15814 & 2 & 1& 56645 & 3 & 1 & 182 \\
 \hline
 \hline
\multicolumn{11}{|l|}{Swing desktop applications}   \\
 \hline
 S3dropbox & 37751 &  0 &  0 & 30609 & 0 & 0 & 115324 & 1 & 1 & 210  \\
 \hline
  SudokuSolver&  27730&  3 &  2 & 20907 & 3 & 2 & 39299 & 2 & 2 & 356  \\
 \hline
 \hline
\multicolumn{11}{|l|}{Android mobile applications}   \\
 \hline
 SGTPuzzler & 13631 / 13865&  1 / 16 &  0 / 0 & 9546 / 9682& 0 / 4& 0 / 1 & 35198 / 35756 & 0 / 1  & 0 / 1& 104 \\
 \hline
 Fennec & 14058 / 14387 &  1 / 1 &  0 / 0 & 8263 / 8898 & 1 / 1 & 0 / 0& 29125/ 31759 & 3 / 3 & 1 / 1& 433 \\
 \hline
 MyTracks & 24036 / 24036 &  161 / 220 & 0 / 0 & 10803 / 13645 & 119 / 119 & 0 / 0 & 39235 / 110977 & 1 / 1 & 0 / 1 & 1192 \\
\hline
\hline
 Total & 199161 / 199724 &  177 / 251 & 5 / 5 & 149441 / 153054& 133 / 137& 6 / 7 & 577821 / 652756& 21 / 22 & 10 / 12 & 6542\\
\hline
\end{tabular}
\end{center}
\vspace{-15pt}
\Caption{{\label{table:results}Experimental results in finding invalid thread access errors
in multithreaded GUI programs. Column ``CG Size'' shows the
call graph size in terms of node number. Column ``\#Warning''
shows the number of warnings issued by our tool. Column ``\#Bug'' shows
the actual bugs found. Columns ``RTA '', ``0-CFA'',
and ``1-CFA'' show the results of using different
call graph construction algorithms. In each cell of the table
for Android applications, a slash ``/'' separates the result of
using standard call graph construction algorithm and our call
graph construction algorithm in Section~\ref{sec:cgcon} (dealing
with reflection calls and adding \annotationnum native method annotations
for SGTPuzzler). The \newbugs errors found in VirgoFTP, FileBunker, ArecaBackup
S3dropbox, and SudokuSolver are previously  unknown. As a comparison, the
results of the ``Requiring Wrappers'' approach (Section~\ref{sec:straightforward}),
are shown at the far right.} }
\end{table*}

@@

We submitted all \newbugs new errors
to the subject developers. As of February 2012, 1 error in S3dropbox has been confirmed,
we have reproduced 4 errors, and 2 errors in FileBunker and ArecaBackup
had not yet been verified by the developers.

@@

%Figures~\ref{fig:swingerror} amd~\ref{fig:pluginerror} show
%two real errors our tool has found. 
Figure~\ref{fig:swingerror}
shows an invalid thread access error our tool found in S3dropbox.
This error happens when the \CodeIn{deleteCurrentBucket}
call on line 63 throws an exception, making it hard to
detect by testing. We reported this error to the S3dropbox developers. Tom Czarniecki,
a key developer of S3dropbox, confirmed this single-GUI-thread
violation. He mentioned that the S3dropbox project uses
certain design patterns to avoid such violations (e.g.,
actions for UI interaction are encapsulated into a \CodeIn{Worker} interface),
but the developers still overlooked  the error our tool found.
Another reason they overlooked this violation is because some GUI frameworks like Swing
do not provide any runtime checks for invalid thread accesses. The Swing
GUI does not exhibit user-visible faults on some erreoneous executions.
However, as clearly stated in the official documentation~\cite{swing}, %Swing is a single-GUI-threaded GUI toolkit and
accessing Swing GUI objects from non-UI threads risks thread interference
or memory-consistency errors.

@@

Figure~\ref{fig:pluginerror} shows an error found in the EclipseRunner
plugin. This error is event-related. It happens when a 
\CodeIn{ResourceChangeEvent} happens, which then invokes the \CodeIn{refresh}
method on line 415 to update the user interface. In EclipseRunner,
the \CodeIn{refresh} method is
called by 6 different methods from the same non-UI thread. Thus, our
tool issues 6 separate warnings to indicate
6 different ways to trigger this error. 5 of the warnings are redundant.

Besides the above two examples, other errors our tool reported are also
subtle to find. For example, our tool found two new errors in SudokuSolver.
One error only happens when
the given Sudoku is unsolvable; and the other one happens
when  JVM fails to launch the %mail composing window of the
user's default mail client.% (after a user clicks the "eMail Me" button).

We found GUI developers have already
used design patterns, runtime checks, and testing to avoid violating
the single-GUI-thread rule. However, due to the huge space of
possible UI interactions, potential
invalid thread access errors still exist in many corner cases and
can be hard to find. 
%Thus, a static analysis as presented in this paper
%would be useful.

@@

We next compare the 3 call graph construction algorithms (RTA, 0-CFA, and 1-CFA)
used in our experiments. As shown in Table~\ref{table:results},  1-CFA found
more errors than the other two algorithms. This
is because RTA and 0-CFA do not consider the calling context when
constructing a call graph, and 
mix the same method calls by different callers into a single node, thus
introducing imprecision. Although the error paths exist in
less precise graphs, the impreicison prevents our algorithm from finding them.
Figure~\ref{fig:ex} illustrates this point.

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.45]{cgexample}
  \vspace*{-5.0ex}\caption {{\label{fig:ex} (a) shows an example code, in which
\CodeIn{safe()} is a Safe UI method, and \CodeIn{uiAcc()} is a UI-accessing method.
(b) shows a less precise call graph built by RTA or 0-CFA, and (c) shows a more
precise call graph built by 1-CFA. Nodes for constructors are omitted for brevity.
Using the less precise call graph, the error
detection algorithm in Figure~\ref{fig:detectalgorithm} will report an invalid method call chain: \CodeIn{entry()}
$\rightarrow$ \CodeIn{f()} $\rightarrow$ \CodeIn{start()} $\rightarrow$ \CodeIn{run()}
$\rightarrow$ \CodeIn{uiAcc()},
but miss the actual error path: \CodeIn{entry()} $\rightarrow$ \CodeIn{g1()} $\rightarrow$
\CodeIn{g2()} $\rightarrow$ \CodeIn{start()} $\rightarrow$ \CodeIn{run()} $\rightarrow$ \CodeIn{uiAcc()}. The
actual error path can only be found when using the more precise call graph.
}}
\end{figure}

@@

The first variant uses DFS for graph traversal by changing
the data structure of $\mathit{worklist}$ on line 7 in Figure~\ref{fig:detectalgorithm} to
a stack, and changing queue operations (\CodeIn{dequeue} and \CodeIn{enqueueAll}
on lines 10 and 21 in Figure~\ref{fig:detectalgorithm}, respectively)
to the corresponding statck operations (\CodeIn{pop} and \CodeIn{pushAll}).
This variant found 11 errors in total and missed 1 error in the FileBunker subject.
The reason is that DFS tends to search deeper into the graph, and the path it returns
may not be the shortest one and is more likely to be an infeasible path in practice.
%For the FileBunker subject, paths
%returned by DFS are much longer than BFS, and do not actually exist.

@@

The second variant uses exhaustive search to find potential errors. This variant
enumerates all non-cyclic paths from all reachable \CodeIn{Thread.start()} nodes
to each UI-accessing node, and then checks whether each
path spawns a new thread and accesses GUI objects without using safe UI methods.
This variant ran out of memory on all subjects. We checked
the explored paths before the algorithm terminated, and found none of them
revealed a real error. This is because the number of
non-cyclic paths in a graph is exponential in the graph size.
For a realistic call graph which often contains thousands of nodes,
it is infeasible to enumerate all paths.
%only a very few amount of paths can be enumerated within a practical amount of time.

@@

\noindent \textbf{\textit{Limitations.}}
Our technique is limited in three aspects. First, it only considers
non-UI threads that are spawned by the UI-thread after the GUI
is initialized, and ignores other possible non-UI threads (quite unusual)
that are created during the pre-initialization GUI work. One way
to remedy this limitation is to design an analysis to identify
those non-UI threads created before a GUI is launched.
Second, like many bug-finding techniques, our technique
is neither sound nor complete. It may miss true positives due
to the graph search strategy, or
issue false positives due to the conservative nature of
a static analysis. Designing better call graph construction algorithms
and graph search strategies may alleviate this limitation.
Third, our tool cannot compute call relationships
for inter-process communication between components. It also
requires users to manually add annotations to characterize
call relationships that involve native methods. This limitation
may lead to false negatives. Investigating the false negative
rate is ongoing work.
